---
title: "Computing One-Year, Two-Year, and Three-Year Cumulative Returns using Monthly Returns"
author: "Suresh L. Paul"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Research question


*Explain the research question. Be concise and explain what your project is about.*




In this project, I compute the one-year cumulative returns, two-year cumulative returns, and three-year cumulative returns using CRSP monthly stock return data between the years 1990 and 2016. 

The formula for cumulative returns is given as follows,
$$
cumret_{(1,t)} = \prod_{n=1}^{t} (1 + r_{n}).
$$

That is, the one-year return is assimilated over 12 months and is given by,
$$
cumret_{(1,12)} = (1 + r_{1}) \times (1 + r_{2}) \times ... \times (1 + r_{11}) \times (1 + r_{12}),
$$


the two-year return is assimilated over 24 months and is given by,
$$
cumret_{(1,24)} = (1 + r_{1}) \times (1 + r_{2}) \times ... \times (1 + r_{23}) \times (1 + r_{24}),
$$


and the three-year return is assimilated over 36 months and is given by,
$$
cumret_{(1,36)} = (1 + r_{1}) \times (1 + r_{2}) \times ... \times (1 + r_{35}) \times (1 + r_{36}).
$$


Also, list all your *R* libraries and *R* verbs/functions that you plan to use and how? Also include the version of *R* and *RStudio* used in your analysis.


***



# Data analysis



The final analysis data is compiled from 4 different datasets - the monthly individual stock return file (*msf.csv*), the header file containing key firm-level characteristics such as SIC codes(*msfhdr.csv*), market stock return file containing all trading dates since the advent of stock trading (*msp500.csv*), and the list of firms in the S&P 500 index (*msp500list.csv*).



## Exploratory data analysis



This subsection documents the initial diagnostic evaluation of all datasets.



### Loading Libraries


First, I start with installing and loading all required *R* libraries. They include,

-   *data.table()*
-   *tidyverse()*, which includes *ggplot2()* and *dplyr()*
-   *lubridate()*
-   *sqldf()*, which is used for many-to-many merge
-   *broom()*
-   *tidyquant()*


Also, the path to datasets, namely *gitpath* and *dirpath*, are initialized.


```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
# load all libraries ...

library("data.table")
library("tidyverse")
library("lubridate")
library("sqldf")
library("broom")
library("tidyquant")
library("plotly")


#-------------------------------------------------------------------------------
# initialize path to files ...
#-------------------------------------------------------------------------------

# Github path to data
gitpath <- "https://raw.githubusercontent.com/sureshlazaruspaul/"

# Local path to files
dirpath <- "C:/data/stock_data/"
```


***



### Importing strategy and Descriptive Summary



All datasets are imported using the *fread()* function from the *data.table* library package. Of the six imported datasets, four (i.e., msf.csv, msfhdr.csv, msp500.csv, and msp500list.csv) are locally available whereas the remaining two (i.e., ffind.csv and 3factors.csv) is imported from online sources.


#### Stock Header File (msfhdr.csv)

##### Import


```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
#-------------------------------------------------------------------------------
# stock data header file
# - contains all firms
# - contains industry codes, firm name, and identifiers
# - contains beginning and ending date of trading
# - create a column of ones(used for merging later on)
#-------------------------------------------------------------------------------

msfhdr <- fread(paste0(dirpath, "msfhdr.csv")) %>%
  select(c("permno", "hsiccd",
           "hcomnam", "begret", "endret")) %>%
  mutate(
    begret = as.Date(begret, "%d%b%Y"),
    endret = as.Date(endret, "%d%b%Y"),
    dummy = 1
  )
```


##### Describe


-   *head()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
head(msfhdr)
```

-   *tail()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
tail(msfhdr)
```

-   *glimpse()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
glimpse(msfhdr)
```

-   *str()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
str(msfhdr)
```

-   *class()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
class(msfhdr)
```

#### Market Stock Return File (msp500.csv)

##### Import

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
#-------------------------------------------------------------------------------
# monthly market return file ...
# - contains returns of S&P 500 firms (equally-weighted, value-weighted)
#-------------------------------------------------------------------------------

msp500 <- fread(paste0(dirpath, "msp500.csv")) %>%
  mutate(
    date = as.Date(caldt, "%d%b%Y")
  )

# get all trading dates from market return files... 
dates <- msp500 %>% select(date) %>%
  mutate(
    dummy = 1
  )
```

##### Describe

-   *head()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
head(msp500)
```

-   *tail()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
tail(msp500)
```

-   *glimpse()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
glimpse(msp500)
```

-   *str()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
str(msp500)
```

-   *class()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
class(msp500)
```

#### List of S&P 500 firms (msp500list.csv)

##### Import

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
#-------------------------------------------------------------------------------
# get s&p 500 list ...
# - identifes which firms are in S&P
#-------------------------------------------------------------------------------

msp500list <- fread(paste0(dirpath, "msp500list.csv")) %>%
  mutate(
    begdt = as.Date(start, "%d%b%Y"),
    enddt = as.Date(ending, "%d%b%Y")
  ) %>%
  select(
    -start, -ending
  )
```

##### Describe

-   *head()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
head(msp500list)
```

-   *tail()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
tail(msp500list)
```

-   *glimpse()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
glimpse(msp500list)
```

-   *str()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
str(msp500list)
```

-   *class()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
class(msp500list)
```

#### Monthly Stock Return File (msf.csv)

##### Import

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
#-------------------------------------------------------------------------------
# get monthly stock returns ...
# - gives stock returns (monthly) of all firms
#-------------------------------------------------------------------------------

msf <- fread(paste0(dirpath, "msf.csv")) %>%
  mutate(
    date = as.Date(date, "%d%b%Y")
  ) %>%
  select(
    permno, date, ret
  )

summary(msf) # diagnose return file

msf <- msf %>%
  filter(
    !is.na(ret) & ret > -0.99
  )

summary(msf) # diagnose return file
```

##### Describe

-   *head()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
head(msf)
```

-   *tail()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
tail(msf)
```

-   *glimpse()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
glimpse(msf)
```

-   *str()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
str(msf)
```

-   *class()*

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
class(msf)
```

##### Descriptive plots

-   histogram of monthly returns (ret)

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
# canvas
canvas <- msf %>%
  filter(
    !is.na(ret)
  ) %>%
  ggplot() +
  theme_classic() +
  xlim(-1,1) +
  labs(
    title = "Distribution of Monthly Returns",
    subtitle = "Sample: S&P 500 firms only",
    caption = paste("From", min(msf$date), 
                    "to", max(msf$date)),
    x = "Monthly Returns",
    y = "Frequency"
  )

# plot
h1 <- canvas +
  geom_histogram(
    aes(
      x=ret
    ),
    bins = 100,
    fill = "red"
  )

h2 <- h1 +
  geom_histogram(
    aes(
      x=ret
    ),
    bins = 250,
    fill = "green"
  )

h2 +
  geom_histogram(
    aes(
      x=ret
    ),
    bins = 1000,
    fill = "blue"
  )
```


-   boxplot of monthly returns (ret)

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
# canvas
canvasa <- msf %>%
  ggplot() + 
  theme_light() +
  ylim(-100, 100) +
  labs(
    title = "Distribution of Monthly Returns",
    subtitle = "Sample: S&P 500 firms only",
    caption = paste("From", min(msf$date), 
                    "to", max(msf$date)),
    x = "",
    y = "Monthly Returns (%)"
  )

canvasa + 
  geom_boxplot(aes(y = ret*100)) + 
  scale_x_discrete() +
  coord_flip()
```


***

#### Variables and observations

From the *glimpse()* function used in the previous section, we can manually ascertain how many variables and observations were there in each dataset. Alternately, the number of observations can be obtained using the *nrow()* function and the number of variables satisfying each criteria can be obtained using a combination of *length()*, *names()*, and *select_if()* functions.


-   Stock Header File (msfhdr.csv)
    -   number of variables: `r length(names(msfhdr))`
        -   number of numeric variables: `r length(names(dplyr::select_if(msfhdr,is.numeric)))`
            -   number of integer variables: `r length(names(dplyr::select_if(msfhdr,is.integer)))`
            -   number of double/float variables: `r length(names(dplyr::select_if(msfhdr,is.double)))`
        -   number of character variables: `r length(names(dplyr::select_if(msfhdr,is.character)))`
        -   number of factor variables: `r length(names(dplyr::select_if(msfhdr,is.factor)))`
            -   number of factors with levels: 0
        -   number of date variables: `r length(names(dplyr::select_if(msfhdr,is.Date)))`
    -   number of observations:  `r nrow(msfhdr)`

-   Market Stock Return File (msp500.csv)
    -   number of variables: `r length(names(msp500))`
        -   number of numeric variables: `r length(names(dplyr::select_if(msp500,is.numeric)))`
            -   number of integer variables: `r length(names(dplyr::select_if(msp500,is.integer)))`
            -   number of double/float variables: `r length(names(dplyr::select_if(msp500,is.double)))`
        -   number of character variables: `r length(names(dplyr::select_if(msp500,is.character)))`
        -   number of factor variables: `r length(names(dplyr::select_if(msp500,is.factor)))`
            -   number of factors with levels: 0
        -   number of date variables: `r length(names(dplyr::select_if(msp500,is.Date)))`
    -   number of observations: `r nrow(msp500)`

-   List of S&P 500 firms (msp500list.csv)
    -   number of variables: `r length(names(msp500list))`
        -   number of numeric variables: `r length(names(dplyr::select_if(msp500list,is.numeric)))`
            -   number of integer variables: `r length(names(dplyr::select_if(msp500list,is.integer)))`
            -   number of double/float variables: `r length(names(dplyr::select_if(msp500list,is.double)))`
        -   number of character variables: `r length(names(dplyr::select_if(msp500list,is.character)))`
        -   number of factor variables: `r length(names(dplyr::select_if(msp500list,is.factor)))`
            -   number of factors with levels: 0
        -   number of date variables: `r length(names(dplyr::select_if(msp500list,is.Date)))`
    -   number of observations: `r nrow(msp500list)`

-   Monthly Stock Return File (msf.csv)
    -   number of variables: `r length(names(msf))`
        -   number of numeric variables: `r length(names(dplyr::select_if(msf,is.numeric)))`
            -   number of integer variables: `r length(names(dplyr::select_if(msf,is.integer)))`
            -   number of double/float variables: `r length(names(dplyr::select_if(msf,is.double)))`
        -   number of character variables: `r length(names(dplyr::select_if(msf,is.character)))`
        -   number of factor variables: `r length(names(dplyr::select_if(msf,is.factor)))`
            -   number of factors with levels: 0
        -   number of date variables: `r length(names(dplyr::select_if(msf,is.Date)))`
    -   number of observations: `r nrow(msf)`

#### Descriptive statistics

Here, the *summary()* and *quantile()* functions are used to obtain statistical summaries on each study variable.


##### Stock Header File (msfhdr.csv)

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
summary(msfhdr)
```


##### Market Stock Return File (msp500.csv)


```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
summary(msp500)
```

##### List of S&P 500 firms (msp500list.csv)


```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
summary(msp500list)
```

##### Monthly Stock Return File (msf.csv)


```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
summary(msf$ret)
quantile(msf$ret, probs = c(0.01, 0.05, 0.10, 0.25, 0.50, 
                            0.75, 0.90, 0.95, 0.99), na.rm = FALSE)
```

### Treatment of missing variables

The data on monthly returns has missing values. All missing values are replaced with zero returns.

## Empirical data analysis

This section details the main empirical analysis.

### Merging, subsetting and filtering

First, we start with the universe of all publicly traded firms in CRSP header file (*msfhdr*). To these firms, we merge with all possible trading dates obtained from market returns file (*msp500*). This merge is a many-to-many merge.


```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
#-------------------------------------------------------------------------------
# for the full list of firms, get all possible trading dates ...
#-------------------------------------------------------------------------------

stock_data1 <- msfhdr %>%
  # get unique firms
  distinct() %>%
  # get trading dates for each firm 
  left_join( # merge dates ...
    dates,
    by = "dummy"
  ) %>%
  select(-dummy)

summary(stock_data1) # diagnose file
glimpse(stock_data1)
```

We join `r nrow(msfhdr)` unique firms with `r nrow(dates)` unique trading days to obtain `r nrow(stock_data1)` observations.

</br>

Second, we obtain monthly returns for all these firms on the respective trading days by merging with the monthly returns file (*msf*). This is a simple left merge. We, then, drop all observations outside starting trading date (*begret*, also known as IPO date) and ending trading date (*endret*, also known as Delisted/Defunct date) for each firm obtained from the header file.


```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
#-------------------------------------------------------------------------------
# for all possible trading dates, see if there is returns ...
#-------------------------------------------------------------------------------

stock_data2 <- stock_data1 %>%
  # get return data
  left_join(
    msf,
    by = c("permno", "date")
  ) %>%
  # limit to trading trades
  filter(
    date >= begret & date <= endret
  ) %>%
  select(-begret, -endret)

summary(stock_data2) # diagnose file

#-------------------------------------------------------------------------------
# Question? 
# Return should be there, but missing. 
# What to do with missing returns?
#-------------------------------------------------------------------------------

glimpse(stock_data2)
```

We start with `r nrow(stock_data1)` observations and end up with `r nrow(stock_data2)` observations after imposing data restrictions.

</br>

Third, we only keep S&P 500 firms. For this purpose, we merge with the file conatining the list of all S&P 500 firms (*msp500list*). This is a simple left merge. We, then, drop all observations outside starting S&P entry date (*begdt*) and ending S&P exit date (*enddt*) for each firm in our dataset.


```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
#-------------------------------------------------------------------------------
# restrict firms to S&P 500 firms only ...
# get s&p list data ...
#-------------------------------------------------------------------------------

stock_data3 <- stock_data2 %>%
  # when was the firm in the s&p list?
  left_join(
    msp500list,
    by = c("permno")
  )  %>%
  # restrict sample
  filter(
    # drop unmerged data (S&P 500 firms only)
    !is.na(begdt) &
      # keep only firms within date range
      (date >= begdt & date <= enddt) &
      # define a sample period, if any
      between(date, as.Date("1925-01-01"), 
                    as.Date("2020-12-31"))
  ) %>%
  select(-begdt, -enddt) %>%
  # what to do with missing returns? 
  # replace with ZERO
  mutate(
    ret = ifelse(is.na(ret), 0, ret)
  ) 

summary(stock_data3) # diagnose file
glimpse(stock_data3)
```

We start with `r nrow(stock_data2)` observations and end up with `r nrow(stock_data3)` observations after imposing data restrictions.

</br>

Finally, we compute the one-year, two-year, and three-year returns by creating 36 new variables - lag 1 of return (*lag1*), lag 2 of return (*lag2*), and so on. We also find the number of missing values for each cumulative returns: from lag 1 (*lag1*) and lag 12 (*lag12*) for one-year return as *nmiss_1yr*, between lag 1 (*lag1*) and lag 24 (*lag24*) for two-year return as *nmiss_2yr*, and between lag 1 (*lag1*) and lag 36 (*lag36*) for three-year return as *nmiss_3yr*. We then use the following formula to compute cumulative returns,
$$
cumret_{(1,t)} = \prod_{n=1}^{t} (1 + r_{n}).
$$


```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
#-------------------------------------------------------------------------------
# How to find return in the previous 12 months? ret1yr
# How to find return in the previous 24 months? ret2yr
# How to find return in the previous 36 months? ret3yr
#-------------------------------------------------------------------------------

annret <- stock_data3 %>%
  mutate(
    lag0 = 1 + ret
  ) %>%
  group_by(permno) %>%
  arrange(permno, date) %>%
  # create lags of return ...
  do(data.frame(., setNames(shift(.$lag0, 1:35), c(paste0("lag", 1:35))))) %>%
  ungroup() %>%
  # count missing
  mutate(
    nmiss_3yr = rowSums(is.na(.))
  ) %>%
  # drop varibles
  select(
    -c(lag24:lag35)
  ) %>%
  # count missing
  mutate(
    nmiss_2yr = rowSums(is.na(.))
  ) %>%
  # drop varibles
  select(
    -c(lag12:lag23)
  ) %>%
  # count missing
  mutate(
    nmiss_1yr = rowSums(is.na(.))
  ) %>%
  # drop varibles
  select(
    -c(lag0:lag11)
  ) %>%
  mutate(
    lag0 = 1 + ret
  ) %>% 
  group_by(permno) %>% 
  arrange(permno, date) %>% 
  # create lags of return ...
  do(data.frame(., setNames(shift(.$lag0, 1:35), c(paste0("lag", 1:35))))) %>%
  ungroup()

rm("dates", "msfhdr", "msp500", "msf", "msp500list",
   "stock_data1", "stock_data2", "stock_data3")

#-------------------------------------------------------------------------------
# initialize all returns to 1
#-------------------------------------------------------------------------------

annret$ret1yr <- 1
annret$ret2yr <- 1
annret$ret3yr <- 1

for(index in 0:11) {
  annret <- annret %>%
    mutate(
      ret1yr = 
        ret1yr * eval(parse(text=paste0("lag", index))))
}

for(index in 0:23) {
  annret <- annret %>%
    mutate(
      ret2yr = 
        ret2yr * eval(parse(text=paste0("lag", index))))
}

for(index in 0:35) {
  annret <- annret %>%
    mutate(
      ret3yr = 
        ret3yr * eval(parse(text=paste0("lag", index))))
}

#-------------------------------------------------------------------------------
# how many missing returns within 12/24/36 months is acceptable?
#   - only 2 per year
#-------------------------------------------------------------------------------

annret <- annret %>%
  mutate(
    ret1yr = ifelse(nmiss_1yr <= 2, ret1yr - 1, NA),
    ret2yr = ifelse(nmiss_2yr <= 4, ret2yr - 1, NA),
    ret3yr = ifelse(nmiss_3yr <= 6, ret3yr - 1, NA)
  ) %>%
  select(
    -ret, -contains("lag", ignore.case = TRUE)
  )

summary(annret) # diagnose file
glimpse(annret)
```

We end up with `r nrow(annret)` observations after imposing data restrictions.

### Univariate Analysis

Here we summarise the cumulative returns.


```{r size="tiny"}
#| echo: FALSE
#| eval: TRUE

min1yr <- round(min(annret$ret1yr, na.rm = TRUE)*100, digits = 3)
min2yr <- round(min(annret$ret2yr, na.rm = TRUE)*100, digits = 3)
min3yr <- round(min(annret$ret3yr, na.rm = TRUE)*100, digits = 3)

max1yr <- round(max(annret$ret1yr, na.rm = TRUE)*100, digits = 3)
max2yr <- round(max(annret$ret2yr, na.rm = TRUE)*100, digits = 3)
max3yr <- round(max(annret$ret3yr, na.rm = TRUE)*100, digits = 3)

mean1yr <- round(mean(annret$ret1yr, na.rm = TRUE)*100, digits = 3)
mean2yr <- round(mean(annret$ret2yr, na.rm = TRUE)*100, digits = 3)
mean3yr <- round(mean(annret$ret3yr, na.rm = TRUE)*100, digits = 3)

sd1yr <- round(sd(annret$ret1yr, na.rm = TRUE)*100, digits = 3)
sd2yr <- round(sd(annret$ret2yr, na.rm = TRUE)*100, digits = 3)
sd3yr <- round(sd(annret$ret3yr, na.rm = TRUE)*100, digits = 3)

p1_1yr <- round(quantile(annret$ret1yr, probs = c(0.01), na.rm = TRUE)*100, digits = 3)
p1_2yr <- round(quantile(annret$ret2yr, probs = c(0.01), na.rm = TRUE)*100, digits = 3)
p1_3yr <- round(quantile(annret$ret3yr, probs = c(0.01), na.rm = TRUE)*100, digits = 3)

p5_1yr <- round(quantile(annret$ret1yr, probs = c(0.05), na.rm = TRUE)*100, digits = 3)
p5_2yr <- round(quantile(annret$ret2yr, probs = c(0.05), na.rm = TRUE)*100, digits = 3)
p5_3yr <- round(quantile(annret$ret3yr, probs = c(0.05), na.rm = TRUE)*100, digits = 3)

p10_1yr <- round(quantile(annret$ret1yr, probs = c(0.10), na.rm = TRUE)*100, digits = 3)
p10_2yr <- round(quantile(annret$ret2yr, probs = c(0.10), na.rm = TRUE)*100, digits = 3)
p10_3yr <- round(quantile(annret$ret3yr, probs = c(0.10), na.rm = TRUE)*100, digits = 3)

q1_1yr <- round(quantile(annret$ret1yr, probs = c(0.25), na.rm = TRUE)*100, digits = 3)
q1_2yr <- round(quantile(annret$ret2yr, probs = c(0.25), na.rm = TRUE)*100, digits = 3)
q1_3yr <- round(quantile(annret$ret3yr, probs = c(0.25), na.rm = TRUE)*100, digits = 3)

q2_1yr <- round(quantile(annret$ret1yr, probs = c(0.50), na.rm = TRUE)*100, digits = 3)
q2_2yr <- round(quantile(annret$ret2yr, probs = c(0.50), na.rm = TRUE)*100, digits = 3)
q2_3yr <- round(quantile(annret$ret3yr, probs = c(0.50), na.rm = TRUE)*100, digits = 3)

q3_1yr <- round(quantile(annret$ret1yr, probs = c(0.75), na.rm = TRUE)*100, digits = 3)
q3_2yr <- round(quantile(annret$ret2yr, probs = c(0.75), na.rm = TRUE)*100, digits = 3)
q3_3yr <- round(quantile(annret$ret3yr, probs = c(0.75), na.rm = TRUE)*100, digits = 3)

p90_1yr <- round(quantile(annret$ret1yr, probs = c(0.90), na.rm = TRUE)*100, digits = 3)
p90_2yr <- round(quantile(annret$ret2yr, probs = c(0.90), na.rm = TRUE)*100, digits = 3)
p90_3yr <- round(quantile(annret$ret3yr, probs = c(0.90), na.rm = TRUE)*100, digits = 3)

p95_1yr <- round(quantile(annret$ret1yr, probs = c(0.95), na.rm = TRUE)*100, digits = 3)
p95_2yr <- round(quantile(annret$ret2yr, probs = c(0.95), na.rm = TRUE)*100, digits = 3)
p95_3yr <- round(quantile(annret$ret3yr, probs = c(0.95), na.rm = TRUE)*100, digits = 3)

p99_1yr <- round(quantile(annret$ret1yr, probs = c(0.99), na.rm = TRUE)*100, digits = 3)
p99_2yr <- round(quantile(annret$ret2yr, probs = c(0.99), na.rm = TRUE)*100, digits = 3)
p99_3yr <- round(quantile(annret$ret3yr, probs = c(0.99), na.rm = TRUE)*100, digits = 3)
```

| Statistic | One-yr Return          | Two-yr Return | Three-yr Return |
|-----------|-----------------------:|--------------:|----------------:|
| Min       | `r min1yr`             | `r min2yr`    | `r min3yr`      |
| P1        | `r p1_1yr`             | `r p1_2yr`    | `r p1_3yr`      |
| P5        | `r p5_1yr`             | `r p5_2yr`    | `r p5_3yr`      |
| P10       | `r p10_1yr`            | `r p10_2yr`   | `r p10_3yr`     |
| Q1 = P25  | `r q1_1yr`             | `r q1_2yr`    | `r q1_3yr`      |
| Mean      | `r mean1yr`            | `r mean2yr`   | `r mean3yr`     |
| Median    | `r q2_1yr`             | `r q2_2yr`    | `r q2_3yr`      |
| Q3 = P75  | `r q3_1yr`             | `r q3_2yr`    | `r q3_3yr`      |
| P90       | `r p90_1yr`            | `r p90_2yr`   | `r p90_3yr`     |
| P95       | `r p95_1yr`            | `r p95_2yr`   | `r p95_3yr`     |
| P99       | `r p99_1yr`            | `r p99_2yr`   | `r p99_3yr`     |
| Max       | `r max1yr`             | `r max2yr`    | `r max3yr`      |
| Std. Dev  | `r sd1yr`              | `r sd2yr`     | `r sd3yr`       |



-   *boxplot()* of one-year return


```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
canvas1 <- annret %>%
  ggplot() + 
  theme_light() +
  ylim(-100, 100) +
  labs(
    title = "Distribution of One-Year Returns",
    subtitle = "Sample: S&P 500 firms only",
    caption = paste("From", min(annret$date), 
                    "to", max(annret$date)),
    x = "",
    y = "One-Year Returns (%)"
  )

canvas1 + 
  geom_boxplot(aes(y = ret1yr*100)) +
  scale_x_discrete() +
  coord_flip()
```

-   *boxplot()* of two-year return

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
canvas2 <- annret %>%
  ggplot() + 
  theme_light() +
  ylim(-100, 100) +
  labs(
    title = "Distribution of Two-Year Returns",
    subtitle = "Sample: S&P 500 firms only",
    caption = paste("From", min(annret$date), 
                    "to", max(annret$date)),
    x = "",
    y = "Two-Year Returns (%)"
  )

canvas2 + 
  geom_boxplot(aes(y = ret2yr*100)) +
  scale_x_discrete() +
  coord_flip()
```



-   *boxplot()* of three-year return


```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
canvas3 <- annret %>%
  ggplot() + 
  theme_light() +
  ylim(-100, 100) +
  labs(
    title = "Distribution of Three-Year Returns",
    subtitle = "Sample: S&P 500 firms only",
    caption = paste("From", min(annret$date), 
                    "to", max(annret$date)),
    x = "",
    y = "Three-Year Returns (%)"
  )

canvas3 + 
  geom_boxplot(aes(y = ret3yr*100)) +
  scale_x_discrete() +
  coord_flip()
```



# Conclusion

To summarise, describe your findings of your empirical analysis. Also, list the total number of *R* verbs that you have used in your study.

# Full *R* code

```{r size="tiny"}
#| echo: TRUE
#| eval: TRUE
library(data.table)
library(tidyverse)
library(lubridate)
library(sqldf)
library(broom)
library(tidyquant)
library(tidyr)

#-------------------------------------------------------------------------------
# path to files ...
#-------------------------------------------------------------------------------

# Github path to data
gitpath <- "https://raw.githubusercontent.com/sureshlazaruspaul/"

# Local path to files
dirpath <- "C:/data/stock_data/"

#-------------------------------------------------------------------------------
# stock data header file
# - contains all firms
# - contains industry codes, firm name, and identifiers
# - contains beginning and ending date of trading
# - create a column of ones(used for merging later on)
#-------------------------------------------------------------------------------

msfhdr <- fread(paste0(dirpath, "msfhdr.csv")) %>%
  select(c("permno", "hsiccd",
           "hcomnam", "begret", "endret")) %>%
  mutate(
    begret = as.Date(begret, "%d%b%Y"),
    endret = as.Date(endret, "%d%b%Y"),
    dummy = 1
  )

head(msfhdr)
glimpse(msfhdr)

#-------------------------------------------------------------------------------
# monthly market return file ...
# - contains returns of S&P 500 firms (equally-weighted, value-weighted)
#-------------------------------------------------------------------------------

msp500 <- fread(paste0(dirpath, "msp500.csv")) %>%
  mutate(
    date = as.Date(caldt, "%d%b%Y")
  )

glimpse(msp500)

# get all trading dates from market return files... 
dates <- msp500 %>% select(date) %>%
  mutate(
    dummy = 1
  )

glimpse(dates)

#-------------------------------------------------------------------------------
# get s&p 500 list ...
# - identifes which firms are in S&P
#-------------------------------------------------------------------------------

msp500list <- fread(paste0(dirpath, "msp500list.csv")) %>%
  mutate(
    begdt = as.Date(start, "%d%b%Y"),
    enddt = as.Date(ending, "%d%b%Y")
  ) %>%
  select(
    -start, -ending
  )

glimpse(msp500list)

#-------------------------------------------------------------------------------
# get monthly stock returns ...
# - gives stock returns (monthly) of all firms
#-------------------------------------------------------------------------------

msf <- fread(paste0(dirpath, "msf.csv")) %>%
  mutate(
    date = as.Date(date, "%d%b%Y")
  ) %>%
  select(
    permno, date, ret
  )

summary(msf) # diagnose return file
glimpse(msf)

msf <- msf %>%
  filter(
    !is.na(ret) & ret > -0.99
  )

summary(msf) # diagnose return file
glimpse(msf)

#-------------------------------------------------------------------------------
# for the full list of firms, get all possible trading dates ...
#-------------------------------------------------------------------------------

stock_data1 <- msfhdr %>%
  # get unique firms
  distinct() %>%
  # get trading dates for each firm 
  left_join( # merge dates ...
    dates,
    by = "dummy"
  ) %>%
  select(-dummy)

summary(stock_data1) # diagnose file
glimpse(stock_data1)

#-------------------------------------------------------------------------------
# for all possible trading dates, see if there is returns ...
#-------------------------------------------------------------------------------

stock_data2 <- stock_data1 %>%
  # get return data
  left_join(
    msf,
    by = c("permno", "date")
  ) %>%
  # limit to trading trades
  filter(
    date >= begret & date <= endret
  ) %>%
  select(-begret, -endret)

summary(stock_data2) # diagnose file

#-------------------------------------------------------------------------------
# Question? 
# Return should be there, but missing. What to do with missing returns?
#-------------------------------------------------------------------------------

glimpse(stock_data2)

#-------------------------------------------------------------------------------
# restrict firms to S&P 500 firms only ...
# get s&p list data ...
#-------------------------------------------------------------------------------

stock_data3 <- stock_data2 %>%
  # when was the firm in the s&p list?
  left_join(
    msp500list,
    by = c("permno")
  )  %>%
  # restrict sample
  filter(
    # drop unmerged data (S&P 500 firms only)
    !is.na(begdt) &
      # keep only firms within date range
      (date >= begdt & date <= enddt) &
      # define a sample period, if any
      between(date, as.Date("1925-01-01"), 
              as.Date("2020-12-31"))
  ) %>%
  select(-begdt, -enddt) %>%
  # what to do with missing returns? 
  # drop missing values (or)
  # filter(
  #  !is.na(ret) & ret >= -0.99
  # ) %>%
  # what to do with missing returns? 
  # replace with ZERO
  mutate(
    ret = ifelse(is.na(ret), 0, ret)
  ) 

summary(stock_data3) # diagnose file
glimpse(stock_data3)

#-------------------------------------------------------------------------------
# DO ALL THE PREVIOUS STEPS IN ONE SHOT
#-------------------------------------------------------------------------------

stock_data <- msfhdr %>%
  # get unique firms
  distinct() %>%
  # get trading dates for each firm 
  left_join( # merge dates ...
    dates,
    by = "dummy"
  ) %>%
  select(-dummy) %>%
  # get return data
  left_join(
    msf,
    by = c("permno", "date")
  ) %>%
  # limit to trading trades
  filter(
    date >= begret & date <= endret
  ) %>%
  select(-begret, -endret) %>%
  # when was the firm in the s&p list?
  left_join(
    msp500list,
    by = c("permno")
  )  %>%
  # restrict sample
  filter(
    # drop unmerged data (S&P 500 firms only)
    !is.na(begdt) &
      # keep only firms within date range
      (date >= begdt & date <= enddt) &
      # define a sample period, if any
      between(date, as.Date("1925-01-01"),
                    as.Date("2020-12-31"))
  ) %>%
  select(-begdt, -enddt) %>%
  # what to do with missing returns?
  # replace with ZERO
  mutate(
    ret = ifelse(is.na(ret), 0, ret)
  ) 

rm("dates", "msfhdr", "msp500", "msf", "msp500list")
rm("stock_data1", "stock_data2", "stock_data3")

summary(stock_data) # diagnose file
glimpse(stock_data)

#-------------------------------------------------------------------------------
# How to find return in the previous 12 months? ret1yr
# How to find return in the previous 24 months? ret2yr
# How to find return in the previous 36 months? ret3yr
#-------------------------------------------------------------------------------

annret <- stock_data %>%
  mutate(
    lag0 = 1 + ret
  ) %>%
  group_by(permno) %>%
  arrange(permno, date) %>%
  # create lags of return ...
  do(data.frame(., setNames(shift(.$lag0, 1:35), c(paste0("lag", 1:35))))) %>%
  ungroup() %>%
  # count missing
  mutate(
    nmiss_3yr = rowSums(is.na(.))
  ) %>%
  # drop varibles
  select(
    -c(lag24:lag35)
  ) %>%
  # count missing
  mutate(
    nmiss_2yr = rowSums(is.na(.))
  ) %>%
  # drop varibles
  select(
    -c(lag12:lag23)
  ) %>%
  # count missing
  mutate(
    nmiss_1yr = rowSums(is.na(.))
  ) %>%
  # drop varibles
  select(
    -c(lag0:lag11)
  ) %>%
  mutate(
    lag0 = 1 + ret
  ) %>% 
  group_by(permno) %>% 
  arrange(permno, date) %>% 
  # create lags of return ...
  do(data.frame(., setNames(shift(.$lag0, 1:35), c(paste0("lag", 1:35))))) %>%
  ungroup()

#-------------------------------------------------------------------------------
# initialize all returns to 1
#-------------------------------------------------------------------------------

annret$ret1yr <- 1
annret$ret2yr <- 1
annret$ret3yr <- 1

for(index in 0:11) {
  annret <- annret %>%
    mutate(
      ret1yr = 
        ret1yr * eval(parse(text=paste0("lag", index))))
}

for(index in 0:23) {
  annret <- annret %>%
    mutate(
      ret2yr = 
        ret2yr * eval(parse(text=paste0("lag", index))))
}

for(index in 0:35) {
  annret <- annret %>%
    mutate(
      ret3yr = 
        ret3yr * eval(parse(text=paste0("lag", index))))
}

#-------------------------------------------------------------------------------
# how many missing returns within 12/24/36 months is acceptable?
#-------------------------------------------------------------------------------

annret <- annret %>%
  mutate(
    ret1yr = ifelse(nmiss_1yr <= 2, ret1yr - 1, NA),
    ret2yr = ifelse(nmiss_2yr <= 4, ret2yr - 1, NA),
    ret3yr = ifelse(nmiss_3yr <= 6, ret3yr - 1, NA)
  ) %>%
  select(
    -ret, -contains("lag", ignore.case = TRUE)
  )



summary(annret) # diagnose file
glimpse(annret)

#-------------------------------------------------------------------------------
# box plot
#-------------------------------------------------------------------------------

canvas1 <- annret %>%
  ggplot() + 
  theme_light() +
  ylim(-100, 100) +
  labs(
    title = "Distribution of One-Year Returns",
    subtitle = "Sample: S&P 500 firms only",
    caption = paste("From", min(annret$date), 
                    "to", max(annret$date)),
    x = "",
    y = "One-Year Returns (%)"
  )

canvas1 + 
  geom_boxplot(aes(y = ret1yr*100)) +
  scale_x_discrete() +
  coord_flip()

#-------------------------------------------------------------------------------

canvas2 <- annret %>%
  ggplot() + 
  theme_light() +
  ylim(-100, 100) +
  labs(
    title = "Distribution of Two-Year Returns",
    subtitle = "Sample: S&P 500 firms only",
    caption = paste("From", min(annret$date), 
                      "to", max(annret$date)),
    x = "",
    y = "Two-Year Returns (%)"
  )
  
canvas2 + 
  geom_boxplot(aes(y = ret2yr*100)) +
  scale_x_discrete() +
  coord_flip()

#-------------------------------------------------------------------------------

canvas3 <- annret %>%
  ggplot() + 
  theme_light() +
  ylim(-100, 100) +
  labs(
    title = "Distribution of Three-Year Returns",
    subtitle = "Sample: S&P 500 firms only",
    caption = paste("From", min(annret$date), 
                      "to", max(annret$date)),
    x = "",
    y = "Three-Year Returns (%)"
  )

canvas3 + 
  geom_boxplot(aes(y = ret3yr*100)) +
  scale_x_discrete() +
  coord_flip()

#-------------------------------------------------------------------------------
# descriptive statistics
#-------------------------------------------------------------------------------

summary(annret$ret1yr)
quantile(annret$ret1yr,
         probs = c(0.01, 0.05, 0.10, 0.25, 0.50, 
                   0.75, 0.90, 0.95, 0.99), na.rm = TRUE)


summary(annret$ret2yr)
quantile(annret$ret2yr,
         probs = c(0.01, 0.05, 0.10, 0.25, 0.50, 
                   0.75, 0.90, 0.95, 0.99), na.rm = TRUE)


summary(annret$ret3yr)
quantile(annret$ret3yr,
         probs = c(0.01, 0.05, 0.10, 0.25, 0.50, 
                   0.75, 0.90, 0.95, 0.99), na.rm = TRUE)
```


\newpage

# Create your CV/Resume here in *R* markdown